A Computational Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2016-10-26 paper/paper.tex
improve the hell out of it

\begin{abstract}
The default mode network (DMN) is the neural representation of the human
baseline mental activity.
%
Many research streams agree on its likely role in evolutionarily adaptive
envisioning of mental scenarios to predict the environment.
The DMN would hence be dedicated to continuous autobiographical memory retrieval,
generation of hypothetical outcomes,
and reward contingency evaluation when letting the mind go.
It explains its highest energy consumption in the brain and
its intimate coupling with conscious awareness.
%
This concept paper proposes a process model that describes
\textit{how} the DMN may actually implement continuous
environmental assessment and prediction.
DMN function is recast in mathematical terms
from the perspective of control theory and
reinforcement learning.
We argue that our formal account of DMN function
explains existing cognitive accounts as special cases and
parsimoneously explains a variety of experimental findings
in animals and humans.
%
Formal models of the computational processes subserved in the DMN
could offer access routes to the statistical regularities
of human predictive behavior.
\end{abstract}

@@

\section{Introduction}
%
When left unperturbed, the human brain is not at rest.
Rather, the brain continues to metabolize large quantities of
oxygen and glucose energy to maintain
potentially purposeful neuronal computation
in the absence of a focused behavioral goal
\citep{kenet2003spontaneously, fiser2004small}.
This baseline energy demand is subject to surprisingly little modulations
when mentally processing environmental challenges
\citep{raichle2001pnas}.
What has early been described as the "stream of consciousness"
in psychology \citep{james1890principles},
found its neurobiological manifestation
in the so-called "default mode network" (DMN)
\citep{shul1997}. This networked collection of some of the highest regions in
the human association cortex \citep{mesulam1998sensation}
consistently
increased in neural activity during
unfocused everyday mind wandering \citep{raichle2001pnas}.
In the beginning of the 21st century, brain imaging was hence the first technology
to allow for the discovery of a unique brain network that subserves
baseline mental activities
\citep{bzdok2015resting}.

The DMN appeared to be exclusive in task-induced
deactivation during various psychological experiments.
This processing of unknown information categories in the DMN has
been argued to mediate an evolutionarily conserved function for the individual.
This is all the more likely because the DMN contains the two
biggest hotspots of energy consumption in the entire central nervous system.
DMN activity also persists to a substantial degree during
sleep and under anesthesia \citep{randy2008}.
Today, many authors believe that the DMN implements
probabilistic estimation of past, hypothetical, and future events. This
brain network
might have emerged to continuously predict environmental events using
mental imagery as an evolutionary advantage.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default mode areas \citep{weiss2006}
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts
\citep{mason2007, seli2016mind}.
%
Understanding of the overarching DMN function is complicated by
simultaneously controlling perception–action cycles and
maintaining baseline contemplations
across time, splace, and content domains
at the interface of the external world and the self.

Consider an agent faced by the choice of the next action
and being guided by rich
reenactment of really happened, hypothetically conceived, and
predicted future events to perfect ongoing behavioral performance.
A particularly attractive computational framework
to describe, quantify, and predict autonomously acting systems like the brain
can be the combination of control theory and reinforcement learning.
It is known that, the more the external world is predictable,
the more mental activity becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation}.
Conversely,
the more the currently executed task is unknown and unpracticed,
the lower the tendency for stimulus-independent thoughts
\citep{filler1973daydreaming, teasdale1995stimulus}.
These ``offline'' processes may however contribute to optimized control of the organism.
We formalize
a policy matrix to capture the space of possible actions that can be performed
on the environment given the current state. A value matrix
maps expected rewards to environmental objects and events.
Switching between states reduces to a sequential processing model.
In a model-based reinforcement learning approach,
we try to understand the environment always better
starting, at birth, with an environment that is completely unknown.
Informed by outcomes of performed action,
the DMN dynamics are constantly adapted in feed-back loops
that are calibrated by prediction error.
A DMN framework incorporating reinforcement learning
can naturally embed human behavior
into the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules.
The implication of the DMN in a diversity of human cognitive processes
can be parsimoneously recast as prediction error minimization
based on pervasive probabilistic simulations,
thus maximizing reward outcome at various time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo sampling
\citep{tenenbaum2011grow}.
\textit{Control} refers to the influences an agent exerts when interacting
with the environment to encourage preferred states.
Even (necessarily imperfect) memories
of past experience, random mind-wandering and dreams during sleep
may provide meaningful building blocks to iteratively improve
the predictive DMN machinery to optimize the behavior of the organism.
%
In short, it has been proposed earlier that
the human brain's energy budget is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
This paper proposes a formal model that
satisfies this contention.

@@

\subsection{Value functions and policies}
The behavior of the agent is governed by some kind of \textit{policy}, which maps states to
sampling probabilities over candidate actions to perform from this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory as follows
\begin{eqnarray*}
  \begin{split}
    S_1 &\sim \rho(.|S_0,A_0), \text{ collect rewards }R_0 = r(S_0, A_0)\\
    A_1 &\sim \pi(.|S_1)\\
    S_2 &\sim \rho(.|S_1,A_1), \text{ collect rewards }R_1 = r(S_1, A_1)\\
    \ldots\\
    A_{t} &\sim \pi(.|S_{t})\\
    S_{t+1} &\sim \rho(.|S_{t},A_{t}), \text{ collect rewards }R_{t} = r(S_{t}, A_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}
Since an action taken now might have repercussions long into the feature, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative version which takes into account the future. A standard for
modeling this accumulation is the so-called
time-discounted \textit{cumulative reward} function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t.
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a
policy $\pi$. The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards.
In the definition of $G^\pi$ above, the constant $\gamma \in [0, 1]$ is the \textit{discount factor}. The choice $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their immediate rewards and has no horizon
for the future. For there to be any learning at all, a technical requirement is that $\gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
Thus the agent is positively biased towards considering rewards which are will
come soon. The reward importance hence decays exponentially
with distance in the future,
given that there is more uncertainty in the farsighted future.
The reward discount factor $\gamma$ can thus be seen as calibrating
the delay discounting behavior of the intelligent agent.

@@

\section{Relation to other cognitive models}
\subsection{Predictive Coding Hypothesis}
The predictive coding and Bayesian brain framework
is a frequently evoked idea related to the default mode function
\citep{bar2007}.
According to this cognitive account,
cortical responses emerge from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
This dynamic intercourse enables inference on the world by reconciling
gaps between fresh sensory input and stored prior information.
Feed-forward sensory experience is constantly calibrated by
top-down modulation at various hierarchical processing levels.
In fact,
back projections probably outnumber by far the existing input projection in the brain
\citep{salin1995corticocortical}.
This can explain the constructive, generative nature of sensory perception
\citep{friston2010free} and
why motor action is intimately related to sensory expectations
\citep{wolpert1995internal, kording2004bayesian}.
At each stage of neural processing,
an internally generated prediction of the sensory input is
directly compared against the actual input from the environment.
A prediction error at one of the processing levels,
i.e. the difference between sensory input
and internally predicted sensation,
incurs plasticity changes of neuronal back-projections (i.e., model parameters)
for gradually improved future prediction of the environment.
Contextual integration is maintained by top-down modulation of sensorimotor
processing by a-priori information about the environment.
The generative model of how data arise in the
world would hence be incorporated into
the current neuronal wiring.
Indeed,
This process permits updates of the internal model of the environment
to best fit the incoming sensory examples.

\begin{itemize}
  \item Both the predictive coding account and MDPs
  are process theories backed up by plausible
  neurobiological evidence that
  view the brain as a ``statistical organ''
  generalizing from the past to new experiences.
  \item Both expose a mechanisms of brain plasticity in that
  neuronal wiring gets increasingly adapted
  when faced by unprecedented environmental
  challenges.
  \item The hierarchical aspects from predictive coding
  is re-expressed in MDPs in form of
  nested prediction of probable upcoming actions and rewards.
  \item Both model the consequences of action. In RL, the horizon of that
  look into the future is manifested in the $\gamma$ parameter
  in the Bellman equation, but not
  explicitly modeled in the PC account that
  mostly emphasizes prediction about the
  immediate future.
  \item Mismatch negativity of PC is immanent in reward prediction error
  by the difference between actually predicted reward of an action given
  a state and the reward predicted by non-linear regression with graded
  discounting of future rewards.
  \item Sensory experience is a generative process from both views.
  In predictive coding, sensory perception of the external world
  is a generative experience due to modulatory top down experience at
  various hierarchical levels of sensory input processing.
  In our RL view, the (partially observed)
  environmental model is incorporated in the Markov decision
  process, \textit{which can be fully recovered based on the last
  state alone}.
  \item Adapting the neuronal connections for improved top-down modulation
  takes the concrete form of gradient computation
  \textit{and back-propagation} in MDPs and RL,
  although this may not be neurobiologically plausible.
  \item Both provide a parsimonious explanation why the
  human brain decreases processing load devoted to incoming information
  when the environment is predictable and allocate increased
  attentional resources when novel stimuli are encountered.
  Similarly, both propose explanations why
  predicting the future is inherently
  linked to information from the past.
\end{itemize}

\subsection{Semantic Hypothesis}
Another frequently proposed account of DMN function revolves
around forming meaningful associations and analogies between
the current experience and
the abstract knowledge derived from past experiences
\citep{bar2007proactive, binder1999conceptual}.
Analogies might naturally tie incoming novel sensory stimuli to
explicit world knowledge extractable from the environment
\citep{bar2009proactive}.
%
Going much beyond language ifself,
semantic analogies can be used to mentally envision non-existent scenarios
that would improve optimal behavior in the environment
by simulating future events.
Such cognitive processes can afford
the internal generation of necessary information
that is not presented in the surrounding environment
by recombining building blocks of
concept knowledge and episodic memories.
Indeed, in aging humans, remembering the past and imaging the future
equally decreased in the level of detail and are associated with
concurrent deficits in forming and integrating relationships between
items \citep{addis2008age, spreng2006temporal}.
A constructive account further explains the reciprocal relationship
between an egocentric first person perspective and
an allocentric bird’s eye perspective immersed in
self-reflection, social knowledge, and autobiographical memories.
%
It is also closely related to episodic memory, language, problem solving,
planning, estimating others' thoughts, and spatial navigation
as these necessitate abstract world knowledge and abstract assocations
for binding the constituent elements in mental scene construction
\citep{schacter2007remembering}.
These scene generation processes could contribute to interpreting the
present and foretelling the future.
This view is for instance supported by evidence in animals that
could learn a \textit{cognitive map of the environment},
even without reward incentives, and exploit it later
for other means \citep{tolman1948cognitive}.

\begin{itemize}
  \item The notion of semantic or knowledge association is
  incorporated into the MDP as the Markov property,
  that is, the current state directly results from the
  agent's history of states and actions. The learned
  value matrix and action transition schedules drive
  stimulus processing and action choice in the present.
  \item In both semantic hypothesis and MDPs,
  there is no evidence to indicate that predictions of various
  levels of complexity, abstraction, timescale and purpose
  use mechanisms that are qualitatively different. This concurs with
  DMN activity increases across time, space, and content domains
  demonstrated in many neuroimaging studies. The semantic hypothesis
  and RL account provide explanations why hippocampus lesion does
  not only impair retrieving memories, but also hypothetical and future
  thinking \citep{hassabis2007patients}.
  \item Both the semantic hypothesis and MDP expose mechanisms of
  how alternative decision trees could be mentally explored.
  \item While mental scene construction implies a distinction between
  environmental perception and internally generated mind-wandering\footnote{
  ``A computational model of how such a process might be structured
  is far from being defined, but it will probably require a form of
  regulation by which perception of the current world is suppressed
  while simulation of possible alternatives are constructed,
  followed by a return to perception of the present.''\citep{buckner2007self}},
  the MDP naturally integrates the former egocentric
  (related to current action, state and reward) and the later
  allocentric (related to past and future actions, states, and rewards)
  angles on the world in a same optimization problem.
  \item The semantic account of DMN function does not offer
  a mechanistic explanation how semantic world knowledge and analogies thereof
  lead to prediction of future actions and states.
  \item In contrast to existing accounts on semantics and
  mental scene construction, the random and creative aspects of DMN function
  are explained in MDPs by the advantages of stochastic optimization.
  \item The semantic hypothesis does not explain why memory recall
  for scene construction in humans is typically fragmentary and noisy
  instead of accurate and reliable.
  Yet, the MDP framework provides an algorithmic explanation in that
  stochasticity of the parameter space search implemented
  by Monte Carlo solvers provably yields better models of the world.
  That is, the purposeful stochasticity of policy and value estimation
  in MDPs provides a candidate explanation for why humans
  have evolved imperfect memories as the better adaptation.
  \item While both semantic and MDP account propose memory-based internally
  generated information for probabilistic mental models of action outcomes,
  only MDPs render explicit the grounds on which the final action is
  eventually chosen (i.e., the estimated cumulative reward).
  \item In sum, episodic scene construction according to the semantic
  hypothesis is lacking an explicit time and incentive model.
\end{itemize}

\subsection{Sentinel Hypothesis}
Processing self-relevant information was perhaps the first
cognitive account that was proposed for the DMN \citep{gusnard2001medial}.
Since then,
different investigators have speculated that neural activity in the DMN
may reflect the brain’s relentless tracking of
relevance in the environment
as an advantageous evolutionary adaptation.
According to this cognitive account, the brain's baseline realizes
a ``radar'' function to
detect subjectively salient and unexpected events that
are the most important for unfolding behavior,
including resources,
the securing of mates, and protection.
%
In fact,
environmental stimuli important for humans are frequently of
social nature. This is probably unsurprising
given that a key property of the human species is
the complexity of their social systems
\citep{tomasello2009cultural}.
More specifically,
according to the ``social brain hypothesis'',
the human brain did not evolve to solve problems of the
physical environment, but to form and maintain increasingly complex
social systems to solve ecological problems by means of social connections
\citep{whiten1988machiavellian}.
Highly efficient neurobiological processing of social cues is exemplified by
facial judgments being routinely processed in less than 100 ms
\citep{bar2006very}.
Our human ancestors were probably among the few organisms that
were more likely to be killed by members of their own species
rather than predators from another species.
Being able to detect the intention and upcoming actions of conspecifics
turned into a decisive evolutionary advantage
\citep{frith2010social}.
%
This may explain why social topics amount to roughly
two thirds of human everyday communication \citep{dunbar1997human}
and
why mind-wandering and dreams
are so rich in stories about people and the complex relationships between them.

 \begin{itemize}
    \item Processing social information has been proposed to underlie the
    physiological baseline of human brain function
    \citep{schilbach2008minds}. This was later challenged by observing
    analogues of the DMN in monkeys \citep{mantini2011default}
    and rats \citep{lu2012rat}, with
    supposedly less advanced social capacities.
    The MDPs parsimoneously explain the dominance of social content in
    human mental scenes by their extremely high relevance to humans
    reflected by high values for social information in the value matrix.
    \item How can a same neurobiological circuit be equally important
    for baseline house-keeping functions and specific task performance?
    Our computational account can explain why the DMN is implicated
    in both a goal-directed task and an idlying rest cognitive set,
    if environmental relevance is processed, manipulated, and retained
    also as a baseline function of human mental activity.
    In tasks, the policy and value matrices are updated to optimize short-term action,
    whereas, at rest, these parameter updates may
    improve especially mid- and long-term action.
    \item The sentinel account does not provide a formal mechanisms of
    how attention and memory resources are exactly reallocated when
    encountering a salient environmental stimulus. In contrast,
    the Bellman recursion is weighted at each step by
    $\sum_{a \in A}\pi(a|s)$. This factor weighs which action cascades
    are recursively explored by the agent and which decision trees are neglected.
    This directly implies allocation of attention and computational load.
    \item In sum,
    MDPs imply social thinking as a physiological brain baseline
    as a special case, without excluding other topics.
 \end{itemize}

>>>>

\begin{abstract}
The default mode network (DMN) is the brain manifestations of the human
baseline mental activity.
%
Many research streams agree on its likely role in evolutionarily adaptive
envisioning of mental scenarios to predict the environment.
The DMN would hence be dedicated to continuous autobiographical memory retrieval,
generation of hypothetical outcomes,
and reward contingency evaluation when letting the mind go.
It explains its highest energy consumption in the brain and
its intimate coupling with conscious awareness.
%
This concept paper proposes a process model that describes
\textit{how} the DMN may actually implement continuous
environmental assessment and prediction to guide action choices.
DMN function is recast in mathematical terms
from the perspective of control theory and
reinforcement learning.
We argue that our formal account of DMN function
explains existing cognitive accounts as special cases and
parsimoneously explains a variety of experimental findings
in animals and humans.
%
Formal models of the computational processes subserved in the DMN
could offer access routes to the statistical regularities
of human predictive behavior.
\end{abstract}

@@

\section{Introduction}
%
When left unperturbed, the human brain is not at rest.
Rather, the brain continues to metabolize large quantities of
oxygen and glucose energy to maintain
potentially purposeful neuronal computation
in the absence of a focused behavioral goal
\citep{kenet2003spontaneously, fiser2004small}.
This baseline energy demand is subject to surprisingly little modulations
when mentally processing environmental challenges
\citep{raichle2001pnas}.
What has early been described as the "stream of consciousness"
in psychology \citep{james1890principles}
found its neurobiological manifestation
in the so-called "default mode network" (DMN)
\citep{shul1997}. This networked collection of some of the highest regions in
the human association cortex \citep{mesulam1998sensation, margulies2016situating}
consistently
increased in neural activity during
unfocused everyday mind wandering \citep{raichle2001pnas}.
In the beginning of the 21st century, brain imaging was hence the first technology
to allow for the discovery of a unique brain network that subserves
baseline mental activities
\citep{bzdok2015resting}.

The DMN appeared to be exclusive in task-induced
deactivation during various psychological experiments.
This processing of unknown information categories in the DMN has
been argued to mediate an evolutionarily conserved function for the individual.
This is all the more likely because the DMN contains the two
biggest hotspots of energy consumption in the entire central nervous system
\citep{raichle2001pnas}.
DMN activity also persists to a substantial degree during
sleep and under anesthesia \citep{randy2008}.
Today, many authors believe that the DMN implements
probabilistic estimation of past, hypothetical, and future events. This
brain network
might have emerged to continuously predict environmental events using
mental imagery as an evolutionary advantage.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default mode areas \citep{weiss2006}
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts
\citep{mason2007, seli2016mind}.
%
Understanding of the overarching DMN function is complicated by
simultaneously controlling perception-action cycles and
maintaining baseline contemplations
across time, splace, and content domains
at the interface of the external world and the self.

Consider an agent faced by the choice of the next action
and being guided by rich
reenactment of really happened, hypothetically conceived, and
predicted future events to perfect ongoing behavioral performance.
A particularly attractive computational framework
to describe, quantify, and predict autonomously acting systems like the brain
can be the combination of control theory and reinforcement learning.
It is known that, the more the external world is predictable,
the more mental activity becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation}.
Conversely,
the more the currently executed task is unknown and unpracticed,
the lower the tendency for stimulus-independent thoughts
\citep{filler1973daydreaming, teasdale1995stimulus}.
These ``offline'' processes may however contribute to optimized control of the organism.
We formalize
a policy matrix to capture the space of possible actions that can be performed
on the environment given the current state. A value matrix
maps expected rewards to environmental objects and events.
Switching between states reduces to a sequential processing model.
In a model-based reinforcement learning approach,
we try to understand the environment always better
starting, at birth, with an environment that is completely unknown.
Informed by outcomes of performed action,
the DMN dynamics are constantly adapted in feed-back loops
that are calibrated by prediction error.
A DMN framework incorporating reinforcement learning
can naturally embed human behavior
into the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules.
The implication of the DMN in a diversity of human cognitive processes
can be parsimoneously recast as prediction error minimization
based on pervasive probabilistic simulations,
thus maximizing reward outcome at various time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo sampling
\citep{tenenbaum2011grow}.
\textit{Control} refers to the influences that an agent exerts when interacting
with the environment to encourage preferred states.
Even (necessarily imperfect) memories
of past experience, random mind-wandering and dreams during sleep
may provide meaningful building blocks to iteratively improve
the predictive DMN machinery to optimize the behavior of the organism.
%
In short, it has been proposed earlier that
the human brain's energy budget is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
This paper proposes a formal model that
satisfies this contention.

@@

\subsection{Value functions and policies}
The behavior of the agent is governed by some kind of \textit{policy}, which maps states to
sampling probabilities over candidate actions to perform from this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory as follows
\begin{eqnarray*}
  \begin{split}
    S_1 &\sim \rho(.|S_0,A_0), \text{ collect rewards }R_0 = r(S_0, A_0)\\
    A_1 &\sim \pi(.|S_1)\\
    S_2 &\sim \rho(.|S_1,A_1), \text{ collect rewards }R_1 = r(S_1, A_1)\\
    \ldots\\
    A_{t} &\sim \pi(.|S_{t})\\
    S_{t+1} &\sim \rho(.|S_{t},A_{t}), \text{ collect rewards }R_{t} = r(S_{t}, A_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}
Since an action taken now might have repercussions long into the feature, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account the future. A standard for
modeling this accumulation is the so-called
time-discounted \textit{cumulative reward} function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t.
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a
policy $\pi$. The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards.
In the definition of $G^\pi$ above, the constant $\gamma \in [0, 1]$ is the \textit{discount factor}. The choice $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their immediate rewards and has no horizon
for the future. For there to be any learning at all, a technical requirement is that $\gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
Thus the agent is positively biased towards considering rewards which are will
come soon. The reward importance hence decays exponentially
with distance in the future,
given that there is more uncertainty in the farsighted future.
The reward discount factor $\gamma$ can thus be seen as calibrating
the delay discounting behavior of the intelligent agent.
It is also important to appreciate that a stochastic policy estimation
is more advantageous in many RL settings.
For instance, a deterministic strategy in playing rock-paper-scissors can
be easily exploited and a uniformly random choice is optimal.

@@

\section{Relation to other cognitive models}
\subsection{Predictive Coding Hypothesis}
The predictive coding and Bayesian brain framework
is a frequently evoked idea related to the default mode function
\citep{bar2007}.
According to this cognitive account,
cortical responses emerge from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
This dynamic intercourse enables inference on the world by reconciling
gaps between fresh sensory input and stored prior information.
Feed-forward sensory experience is constantly calibrated by
top-down modulation at various hierarchical processing levels.
In fact, axonal
back projections probably outnumber by far the input projection
existing in the brain \citep{salin1995corticocortical}.
This can explain the constructive, generative nature of sensory perception
\citep{friston2010free} and
why motor action is intimately related to sensory expectations
\citep{wolpert1995internal, kording2004bayesian}.
At each stage of neural processing,
an internally generated prediction of the sensory input is
directly compared against the actual input from the environment.
A prediction error at one of the processing levels,
i.e. the difference between sensory input
and internally predicted sensation,
incurs plasticity changes of neuronal back projections (i.e., model parameters)
for gradually improved future prediction of the environment.
Contextual integration is hence maintained by top-down modulation of sensorimotor
processing by a-priori information about the environment.
The generative model of how perceived sensory events arise in the
world would be incorporated into
the current neuronal wiring.
Indeed,
This process permits updates of the internal model of the environment
to best fit the constant influx of sensory examples.

\begin{itemize}
  \item Both the predictive coding account and MDPs
  are process theories backed up by plausible
  neurobiological evidence that
  view the brain as a ``statistical organ''
  generalizing from the past to new experiences.
  \item Both provide a parsimonious explanation why the
  human brain decreases processing load devoted to incoming information
  when the environment is predictable and allocate increased
  attentional resources when novel stimuli are encountered.
  Similarly, both propose explanations why
  predicting the future is inherently
  linked to information from the past.
  \item Sensory experience is a generative process from both views.
  In predictive coding, sensory perception of the external world
  is a generative experience due to modulatory top down experience at
  various hierarchical levels of sensory input processing.
  In our RL view, the (partially observed)
  environmental model is incorporated in the MDP,
  \textit{which can be fully recovered based on the last
  state alone}.
  \item Both naturally expose a mechanisms of brain plasticity in that
  neuronal wiring gets increasingly adapted
  when faced by unprecedented environmental
  challenges.
  \item The hierarchical aspects from predictive coding
  is re-expressed in MDPs in form of
  nested prediction of probable upcoming actions and rewards.
  \item Both model the consequences of action. In RL, the horizon of that
  look into the future is explicitly manifested in the $\gamma$ parameter
  in the Bellman equation, but not
  explicitly modeled in the PC account that
  tends to emphasize prediction about the
  immediate future.
  \item Mismatch negativity of PC is immanent in reward prediction error
  by the difference between actually predicted reward of an action given
  a state and the reward predicted by non-linear regression with graded
  discounting of future rewards.
  \item Adapting the neuronal connections for improved top-down modulation
  takes the concrete form of gradient computation
  \textit{and back-propagation} in MDPs and RL,
  although this may not be neurobiologically plausible.
  \item In sum,
  the PC account in cognitive neuroscience does not expose an explicit
  mechanism of the horizon of future considerations or
  how the internal model of the world gets updated,
  and collapses sensory input processing and action output preparation.
\end{itemize}

\subsection{Semantic Hypothesis}
Another frequently proposed cognitive account of DMN function revolves
around forming logical associations and analogies between
the current experience and
the conceptual knowledge derived from past experiences
\citep{bar2007proactive, binder1999conceptual}.
Analogies might naturally tie incoming novel sensory stimuli to
explicit world knowledge (i.e., semantics) extractable from the environment
\citep{bar2009proactive}.
%
Going much beyond human language ifself,
semantic building blocks provide the basis to mentally envision non-existent scenarios
that would improve optimal behavior in the environment
by simulating future events.
Such cognitive processes can afford
the internal generation of necessary information
that is not presented in the surrounding environment
by recombining building blocks of
concept knowledge and episodic memories.
Indeed, in aging humans, remembering the past and imaging the future
equally decreased in the level of detail and are associated with
concurrent deficits in forming and integrating relationships between
items \citep{addis2008age, spreng2006temporal}.
Further,
a constructive account explains the reciprocal relationship
between an egocentric first person perspective and
an allocentric bird’s eye perspective immersed in
self-reflection, social knowledge, and autobiographical memories.
%
Cognitive aspects of egocentric-allocentric switching
are also closely related to episodic memory, language, problem solving,
planning, estimating other people's thoughts, and spatial navigation
as these necessitate abstract world knowledge and abstract assocations
for binding the constituent elements in mental scene construction
\citep{schacter2007remembering}.
These scene generation processes could contribute to interpreting the
present and foretelling the future.
This view is for instance supported by evidence in animals that
could learn a \textit{cognitive map of the environment},
even without reward incentives, and exploit it later
for other means \citep{tolman1948cognitive}.

\begin{itemize}
  \item Both the semantic hypothesis and MDPs expose mechanisms of
  how alternative decision trees could be mentally explored.
  \item In both semantic hypothesis and MDPs,
  there is no evidence to indicate that predictions of various
  levels of complexity, abstraction, timescale and purpose
  use mechanisms that are qualitatively different. This concurs with
  DMN activity increases across time, space, and content domains
  demonstrated in many neuroimaging studies. The semantic hypothesis
  and RL account provide explanations why hippocampus lesion does
  not only impair retrieving memories, but also hypothetical and future
  thinking \citep{hassabis2007patients}.
  \item The notion of semantic or knowledge association is
  incorporated into the MDP as the Markov property,
  that is, the current state directly results from the
  agent's history of states and actions. The learned
  value matrix and action transition schedules drive
  stimulus processing and action choice in the present.
  \item While mental scene construction implies a distinction between
  environmental perception and internally generated mind-wandering\footnote{
  ``A computational model of how such a process might be structured
  is far from being defined, but it will probably require a form of
  regulation by which perception of the current world is suppressed
  while simulation of possible alternatives are constructed,
  followed by a return to perception of the present.''\citep{buckner2007self}},
  the MDPs naturally integrate the former egocentric
  (more related to current action, state and reward) and the later
  allocentric (more related to past and future actions, states, and rewards)
  angles on the world in a same optimization problem.
  \item The semantic account of DMN function does not offer
  a mechanistic explanation how explicit world knowledge and semnatic analogies thereof
  lead to prediction of future actions and states.
  \item In contrast to existing accounts on semantics and
  mental scene construction, the random and creative aspects of DMN function
  are explained in MDPs by the advantages of stochastic optimization.
  \item The semantic hypothesis does not explain why memory recall
  for scene construction in humans is typically fragmentary and noisy
  instead of accurate and reliable.
  Yet, the MDP framework provides an algorithmic explanation in that
  stochasticity of the parameter space search implemented
  by Monte Carlo solvers provably yields better models of the world.
  That is, the purposeful stochasticity of policy and value estimation
  in MDPs provides a candidate explanation for why humans
  have evolved imperfect noisy memories
  as the more advantageous adaptation.
  \item While both semantic and MDP account propose memory-based internally
  generated information for probabilistic mental models of action outcomes,
  only MDPs render explicit the grounds on which the final action is
  eventually chosen (i.e., the estimated cumulative reward).
  \item In sum, episodic scene construction according to the semantic
  hypothesis is lacking an explicit time and incentive model;
  neither does it explain how randomness of human mental experience
  can be beneficial.
\end{itemize}

\subsection{Sentinel Hypothesis}
Processing self-relevant information was perhaps the first
cognitive account that was proposed for the DMN \citep{gusnard2001medial}.
Since then,
different investigators have speculated that neural activity in the DMN
may reflect the brain’s relentless tracking of
relevance in the environment
as an advantageous evolutionary adaptation.
According to this cognitive account, the brain's baseline realizes
a ``radar'' function to
detect subjectively salient and unexpected events that
are the most important for unfolding behavior,
including resources,
the securing of mates, and protection.
%
In fact,
environmental stimuli important for humans are frequently of
social nature. This is probably unsurprising
given that a key property of the human species is
the complexity of their social systems
\citep{tomasello2009cultural}.
More specifically,
according to the ``social brain hypothesis'',
the human brain did not evolve to solve problems of the
physical environment, but to form and maintain increasingly complex
social systems to solve ecological problems by means of social relationships
\citep{whiten1988machiavellian}.
Highly efficient neurobiological processing of social cues is exemplified by
facial judgments being routinely processed in less than 100 ms
\citep{bar2006very}.
Our human ancestors were probably among the few organisms that
were more likely to be killed by members of their own species
rather than predators from another species.
Being able to detect the intention and upcoming actions of conspecifics
turned into a decisive evolutionary advantage
\citep{frith2010social}.
%
This may explain why social topics amount to roughly
two thirds of human everyday communication \citep{dunbar1997human}
and
why mind-wandering and dreams
are so rich in stories about people and the complex relationships between them.

 \begin{itemize}
    \item Processing social information has been proposed to underlie the
    physiological baseline of human brain function
    \citep{schilbach2008minds}. This was later challenged by observing
    analogues of the DMN in monkeys \citep{mantini2011default}
    and rats \citep{lu2012rat}, with
    supposedly less advanced social capacities.
    The MDPs parsimoneously explain the dominance of social content in
    human mental scenes by their extremely high relevance to humans
    reflected by high values for social information in the value matrix.
    \item How can a same neurobiological circuit be equally important
    for baseline house-keeping functions and specific task performance?
    Our computational account can explain why the DMN is implicated
    in both a goal-directed task and an idlying rest cognitive set,
    if environmental relevance is processed, manipulated, and retained
    also as a baseline function of human mental activity.
    In tasks, the policy and value matrices are updated to optimize short-term action,
    whereas, at rest, these parameter updates may
    improve especially mid- and long-term action.
    \item The sentinel account does not provide a formal mechanisms of
    how attention and memory resources are exactly reallocated when
    encountering a salient environmental stimulus. In contrast,
    the Bellman recursion is weighted at each step by
    $\sum_{a \in A}\pi(a|s)$. This factor weighs which action cascades
    are recursively explored by the agent and which decision trees are neglected.
    This directly implies allocation of attention and computational load.
    \item In sum,
    MDPs imply a ``radar'' function of montoring the environment for salient information
    in general and explain social thinking as a physiological brain baseline
    as an important special case, without excluding other topics
    contemplated in continuous mental activity.
 \end{itemize}

\section{Conclusion}
What brain function could be important enough
for the existence and survival of the human species
to warrant constant, high energy costs?
MDPs provide an attractive
computational model how the human association cortex
might implement supra-modal representation and control of the environment to
optimize the organism's fate.
% by obtaining reward and avoiding pain.
This idealized process model explains
a number of previous experimental observations in the
DMN by a simple but non-trivial mechanism.
%
Viewed as an MDP, behavior unfolds by integrating happened past events
and possible future events to guide action choice in the present context.
This functional account is more compatible with the DMN's
poorly understood involvement across
autobiographical memory retrieval, problem solving,
abstract reasoning based on internally generated scenes, social cognition,
as well as delay discounting and self-related prospection.
MDPs provide a mathematical formalism how
optimal substructure in the environment can be recursively exploited
when confronted with complicated decisions.
Algorithmic gains by injecting stochasticity in the recall of past
actions and outcomes to improve the internal world model may explain why
very accurate memories have been disfavored in human evolution
and why human creativity might be adaptive.
%
In principle,
neuroscientific experiments can be designed that operationalize
the set of action, value, and state variables that determine
the behavior of intelligent RL systems.
The proposed machine-learning
approach to DMN function is hence not only practically computable but
also neuroscientifically falsifiable.
At the least, we propose an alternative vocabulary to
describe and interpret neuroscientific studies in DMN research.
%
Ultimately,
the DMN can be viewed as a smoothing kernel extending
from the relatively recent events to anticipated upcoming events
to constantly improve sensory processing and
action performance in the world.
